### 2023-12-12 thesis/sections/5.1-object-identification.tex
results
experiments

The model, the \textbf{object identifier} is split into three submodules, the \emph{bounding box encoder}, the \emph{attribute encoder} and the \emph{identifier}.
Each normalized bounding box of the sample is encoded with the \emph{bounding box encoder}.
Hereby, they are passed through \emph{ResNet-avg}.
The preceding layers \emph{ResNet-1} to \emph{ResNet-4} couldn't be compared due to memory restrictions on the GPU.
A deeper analysis of the effects of different feature extractors will be done in the following chapters.
The resulting vector is flattened and projected to the embedding dimension $e$ via a linear layer.
% SD: This is quite low? Why only 10?
% DK: different embedding sizes are tested (done)
Correspondingly, the one-hot vector of the attributes is projected to the same embedding dimension with another linear layer, using the \emph{attribute encoder}.
In the \emph{identifier} submodule, each bounding box is combined with the representation of the attributes.
By bringing both vectors to the same embedding size, the dot product between each embedded bounding box and the embedded attributes can be calculated.
A high correlation between the attributes and the object should result in a high dot product, while a low correlation results in a lower dot product.
The model can so learn to connect the object representation with the attributes.
The dot products are concatenated and form a vector with as many dimensions as objects present.
The $softmax$ function is applied over the resulting vector, which returns a probability distribution over all objects; the model points to the object with the highest probability.
% The model then points to an object, using the \emph{softmax} function over this resulting vector.

The experiments are conducted with the following hyperparameters: a learning rate of $2\times10^{-4}$, a batch size of 32 samples and 30 epochs, \emph{Adam} \citep{Kingma2015} as optimizer.
The number of epochs was chosen manually after identifying when the loss as well as the test accuracies didn't improve significantly.
8000 randomly selected samples are used for training, the remaining 2000 samples for testing.
The loss is calculated using cross entropy.
The following embedding dimension $e$ are compared: 2, 10, 50, 100, 500, 1000.
% SD: So, the output of this model is a vector of 0 and 1s, with the 1 identifying the corrct object (or the other way around?).
% DK: the output is a probability distriution over all objects (inserted above; done)

>>>>

The model, the \textbf{object identifier} is split into three submodules, the \emph{bounding box encoder}, the \emph{attribute encoder} and the \emph{identifier}.
Each normalized bounding box of the sample is encoded with the \emph{bounding box encoder}.
Hereby, they are passed through \emph{ResNet-avg}.
The preceding layers \emph{ResNet-1} to \emph{ResNet-4} couldn't be compared due to memory restrictions on the GPU.
The resulting vector is flattened and projected to the embedding dimension $e$ via a linear layer.
% SD: This is quite low? Why only 10?
% DK: different embedding sizes are tested (done)
Correspondingly, the one-hot vector of the attributes is projected to the same embedding dimension with another linear layer, using the \emph{attribute encoder}.
In the \emph{identifier} submodule, each bounding box is combined with the representation of the attributes.
By bringing both vectors to the same embedding size, the dot product between each embedded bounding box and the embedded attributes can be calculated.
A high correlation between the attributes and the object should result in a high dot product, while a low correlation results in a lower dot product.
The model can so learn to connect the object representation with the attributes.
The dot products are concatenated and form a vector with as many dimensions as objects present.
The $softmax$ function is applied over the resulting vector, which returns a probability distribution over all objects; the model points to the object with the highest probability.
% The model then points to an object, using the \emph{softmax} function over this resulting vector.

The experiments are conducted with the following hyperparameters: a learning rate of $2\times10^{-4}$ and 30 epochs.
The number of epochs was chosen manually after identifying when the loss as well as the test accuracies didn't improve significantly.
The loss is calculated using cross entropy.
The following embedding dimension $e$ are compared: 2, 10, 50, 100, 500, 1000.
% SD: So, the output of this model is a vector of 0 and 1s, with the 1 identifying the corrct object (or the other way around?).
% DK: the output is a probability distriution over all objects (inserted above; done)

### 2023-12-12 thesis/sections/5.2-re-generation.tex
experiments

For both setups, the same hyperparameters as in the previous experiments are used: a learning rate of $2\times10^{-4}$, a batch size of 32 samples and 30 epochs, \emph{Adam} \citep{Kingma2015} as optimizer.
8000 randomly selected samples are used for training, the remaining 2000 samples for testing.
The loss is calculated using cross entropy.
Table \ref{tab:variables-reference-expression-generation} shows which variables are compared for each model:

>>>>

For both setups, the same hyperparameters as in the previous experiments are used: a learning rate of $2\times10^{-4}$ and 30 epochs.
The loss is calculated using cross entropy.
Table \ref{tab:variables-reference-expression-generation} shows which variables are compared for each model:

### 2023-12-12 thesis/sections/5.3-reference-resolution.tex
experiments

For all setups, the same hyperparameters as in the previous experiments are used: a learning rate of $1\times10^{-4}$, a batch size of 32 samples and 30 epochs, \emph{Adam} \citep{Kingma2015} as optimizer.
8000 randomly selected samples are used for training, the remaining 2000 samples for testing.
The loss is calculated using cross entropy.
Table \ref{tab:variables-reference-resolution} shows the variables that are changed during the experiments for each of the models.

@@

To do this, the model combines a representation of the image and a representation of the target object.
More specifically, the image is encoded using the \emph{image encoder} submodule, omitting the last \emph{max pool} layer.
By this, the resulting matrix has 128\times14\times14 dimension, corresponding to the 14\time14 regions.
For the target object, the incremental RE is passed through an LSTM as described before.
Both encodings are passed through two linear layers separately to project them to the projection dimension $p$, followed by the $tanh$ function.

>>>>

For all setups, the same hyperparameters as in the previous experiments are used: a learning rate of $1\times10^{-4}$ and 30 epochs.
The loss is calculated using cross entropy.
Table \ref{tab:variables-reference-resolution} shows the variables that are changed during the experiments for each of the models.

@@

To do this, the model combines a representation of the image and a representation of the target object.
More specifically, the image is encoded using the \emph{image encoder} submodule, omitting the last \emph{max pool} layer.
By this, the resulting matrix has 128\times14\times14 dimension, corresponding to the 14\time14 regions.
For the target object, the incremental RE is passed through an LSTM as described before.
Both encodings are passed through two linear layers separately to project them to the projection dimension $p$, followed by the $tanh$ function.
Following, the dot product is calculated between the projected image and the projected attributes which results in a 14\times14 matrix.
This emphasizes the correlation between the attributes and each region.
A high dot product for a region indicates a high correlation between the attributes and the specific region, while a low dot product indicates the opposite.
Training a model like this should therefore highlight the regions in the image that are described by the referring expression, namely the region around the target object.
To calculate the loss, the $softmax$ function is applied over the prediction and compared to the ground truth matrix $A$ using the \emph{binary cross entropy}.
Hereby, the models converged faster and were trained for 20 epochs with a learning rate of $2\times10^{-4}$.
An $LSTM_e=15$ and $LSTM_o=1500$ showed the best results.
Projection sizes $p=\{50,100,300,500,1000\}$ are compared.

The \emph{attention reference resolver} is evaluated on the summed predicted probability for the regions in the target area.
In particular, the predicted matrix, consisting of probabilities for each region is multiplied with the ground truth matrix $A$, consisting of only ones and zeros.
The result is summed and returns the probability mass for the target area.
If the model predicts the target area perfectly, the probabilities in the target area sum to 1.
If the model for instance focuses on the wrong object, the probability mass in the target area is lower.

### 2023-12-12 thesis/sections/6-grounding-language-games.tex
experiments

As the experiments before, three different tasks are set up.
For the \emph{object identification} task, the agents play discrimination games similar to the setup in \citep{Lazaridou2017}.
For the \emph{RE generation} task, the receiver needs to generate a human language referring expression based on the emerged language referring expression by the sender.
Since, it is possible that the agents simply learn to align these two languages, the agents are tasked to point towards the center coordinates of the target object in the final task, the \emph{reference resolution} task.

As discussed in Section \ref{sec:background:lg:setup}, language games are evaluated on the success of communication and the emergence of a language.
The testing phase is not necessarily used to test the generalization capabilities of the agents, but more to extract a sample of the emerged language.
It is therefore not as important that the samples in the testing phase are unseen samples.
Hence, for both training and testing a certain number of samples are randomly drawn from the complete dataset.
In these experiments, the dataset consists of 10.000 scenes in total.
To train the agents, 12.800 games are played (using 12.800 randomly drawn samples from the dataset) with a batch size of 32 samples.
To retrieve the samples of the emerged language, after the agents are trained, they are frozen and play 2.048 games with randomly drawn samples from the dataset.
This yields 2.048 interactions that will be used to analyze the emerged language.

>>>>

As the experiments before, three different tasks are set up.
For the \emph{object identification} task, the agents play discrimination games similar to the setup in \citep{Lazaridou2017}.
For the \emph{RE generation} task, the receiver needs to generate a human language referring expression based on the emerged language referring expression by the sender.
Since, it is possible that the agents simply learn to align these two languages, the agents are tasked to point towards the center coordinates of the target object in the final task, the \emph{reference resolution} task.
For all three setups, the sender uses the same setup and only the receiver is switched.
By this, differences in the produced referring expression in the message by the sender are only dependent on the receiver's ability to solve the different tasks.
In particular, the impact of the task on the emerged language can be analyzed in more detail.

As discussed in Section \ref{sec:background:lg:setup}, language games are evaluated on the success of communication and the emergence of a language.
The testing phase is not necessarily used to test the generalization capabilities of the agents, but rather to extract a sample of the emerged language.
It is therefore not as important that the samples in the testing phase are unseen samples.
Hence, for both training and testing a certain number of samples are randomly drawn from the complete dataset.
In these experiments, the dataset consists of 10.000 scenes in total.
To train the agents, 128.000 games are played (using 128.000 randomly drawn samples from the dataset) with a batch size of 32 samples.
To retrieve the samples of the emerged language, after the agents are trained, they are frozen and play 2.048 games with randomly drawn samples from the dataset.
This yields 2.048 interactions that will be used to analyze the emerged language.

### 2023-12-12 thesis/sections/6.1-object-identification.tex
experiments

The discrimination games in this research have a very similar setup as described in \citep{Lazaridou2017}.
The agents in this research resemble their \emph{agnostic sender} as well as their \emph{receiver}.
One central difference is the production of the message.
The main goal of their language game was the identification of the concept that and image was related to.
Therefore, the sender communicated only single-symbol messages to the receiver, which should describe the concept of the target image.
Opposed to that in this research, the agents are tasked to discriminate objects from each other based on their attributes.
% SD: On the contrary,...
% DK: TODO
It is therefore assumed that the sender will communicate these discriminative attributes.
For that reason, the sender is allowed to generate sequences as a message.

@@

The experiments are conducted with the following hyperparameters: a learning rate of $2\times10^{-4}$, a temperature for the Gumbel-Softmax relaxation of 1 and \emph{Adam} \citep{Kingma2015} as optimizer.
The loss is calculated using cross entropy.
The following values for the variables are compared:
\begin{itemize}
    \item $h_s$: 100, 500, 1000
    \item $h_r$: 10, 30, 50, 100, 500, 1000
    \item $|V|$: 1, 10, 16, 50, 100
    \item $n$: 1, 2, 3, 4, 5, 6
\end{itemize}
The language game is run on the 'Dale-2', 'Dale-5' and 'CLEVR color' datasets.

>>>>

The discrimination games in this research have a very similar setup as described in \citep{Lazaridou2017}.
The agents in this research resemble their \emph{agnostic sender} as well as their \emph{receiver}.
One central difference is the production of the message.
The main goal of their language games was the identification of the concept that the shown image was related to.
Therefore, the sender communicated only single-symbol messages to the receiver which should describe the concept of the target image.
Opposed to that, in this research, the agents are tasked to discriminate objects from each other based on their attributes.
% SD: On the contrary,...
% DK: (done)
It is therefore assumed that the sender will communicate these discriminative attributes.
For that reason, the sender is allowed to generate sequences as a message where for instance each symbol in the sequence might correspond to on attribute.

@@

The experiments are conducted with a learning rate of $2\times10^{-4}$, the loss is calculated using cross entropy.
The following values for the variables are compared:
\begin{itemize}
    \item $h_s$: 100, 500, 1000
    \item $h_r$: 10, 30, 50, 100, 500, 1000
    \item $|V|$: 1, 10, 16, 50, 100
    \item $n$: 1, 2, 3, 4, 5, 6
\end{itemize}
The language game is run on the 'Dale-2', 'Dale-5' and 'CLEVR color' datasets.
